Supplementary Information: Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments

### 2025-06-27 doc/cover_letter.Rmd
updated cover letter

April 22, 2025
\newline
\newline

@@

We are pleased to submit our manuscript "Tropical tree drought sensitivity is jointly shaped by drought characteristics, species adaptations, and individual microenvironments" for consideration for publication in *Ecology Letters*.

@@

Here, we use dendrometer band measurements from 1820 trees across 30 representative species in a tropical seasonal forest in northern Thailand to show that tree growth responses to drought can be influenced by drought characteristics, microenvironments and species traits. 
In two ENSO drought years within a 14-year timeseries, species and trees showed heterogeneous drought sensitivity across drought events analysed, as opposed to consistent or correlated responses scaled with drought intensity.
Exposed trees consistently suffered greater growth declines, but the effects of individual and species drivers related to tree-water relations - topographic wetness, tree size and leaf deciduousness - were all heterogenous between a dry- and wet-season drought. 
Our findings imply that although drought resistance is often regarded as a static trait, the species and tree level responses show a rich and dynamic diversity of responses across drought events. 
In sum, our findings of heterogenous responses indicate potential spatiotemporal insurance effects in diverse forests that may ameliorate effects of intensifying drought. 
This is a significant finding that will be of interest to the broad readership of *Ecology Letters*.

We believe our manuscript to be an appropriate length for *Ecology Letters*, with 4989 words and 6 display items in the main article. The article is accompanied by a Supplementary Information file (12 figures).

>>>>

June 27, 2025
\newline
\newline

@@

We are pleased to submit our manuscript “Drought response is not a species trait: tropical tree drought sensitivity is shaped by drought characteristics, species adaptations, and individual microenvironments” for consideration for publication in
*Ecology Letters*.

@@

Here, we use dendrometer band measurements from 1820 trees across 30 representative species in a tropical seasonal forest in northern Thailand to show that tree growth responses to drought can be influenced by drought characteristics, microenvironments and species traits. 
In three drought years within a 14-year timeseries, species and trees showed heterogeneous drought sensitivity across drought events analysed, as opposed to consistent or correlated responses scaled with drought intensity. 
While exposed trees suffered greater growth declines, the effects of individual and species drivers related to tree-water relations - topographic wetness, tree size and leaf deciduousness - were all heterogenous between droughts that differed in climatic characteristics. 
Our findings imply that although drought resistance is often regarded as a static trait, the species and tree level responses show a rich and dynamic diversity of responses across drought events. 
In sum, our findings of heterogenous responses indicate potential spatiotemporal insurance effects in diverse forests that may ameliorate effects of intensifying drought. This is a significant finding that will be of interest to the broad readership of *Ecology Letters*.

We believe our manuscript to be an appropriate length for *Ecology Letters*, with 5019 words and 6 display items in the main article. The article is accompanied by a Supplementary Information file (16 figures).
